### 2024-01-21 analysis.qmd
First push

---
  title: "Sociological Science Categorical Tolerance Follow-up"
  author: "Omar Lizardo"
  date: "`r Sys.Date()`"
---

# Setup
```{r setup, include=FALSE}
     knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
     library(boot)
     library(conflicted)
     library(here)
     library(googlesheets4)
     library(ggplot2)
     library(pscl)
     library(sjPlot)
     library(tidyverse)
     library(webshot)
     source(here("Functions", "dat.demog.R"))
     source(here("Functions", "dat.genres.R"))
```

# Data Wrangling
```{r Getting and wrangling data}
    gs4_deauth()
    dat <- read_sheet("https://docs.google.com/spreadsheets/d/1r6IcGtlTOaAFHxJYgnuGM6IEiJhaIoDBhxEfRIhZEeI/edit?usp=sharing") 
    dat <- dat[-c(1), ] #getting rid of first row
    #working.dat <- dat.wrang(dat)
```

```{r Data Wrangling Genres}
    dat.dis <- dat.genres(dat)$dat.g %>%
          mutate(across(Classical:Metal, function(x) {if_else(x < 3, 1, 0)})) %>% 
          rowwise() %>% 
          mutate(num.dis = sum(c_across(Classical:Metal), na.rm = T)) %>% 
          ungroup()
    head(dat.dis)
         reg.dat <- dat.demog(dat) %>% 
          left_join(dat.dis)
```

```{r Univariate Distribution of Dislikes Plot}
     d <- round(table(dat.dis$num.dis)/nrow(dat.dis), 2)[1:11]
     plot.dat <- data.frame(cbind(count = 0:10, prop = d))
     p <- ggplot(data = plot.dat, aes(x = count, y = prop)) 
     p <- p + geom_line(color = "blue")
     p <- p + geom_point(size = 4, color = "blue")
     p <- p + scale_x_continuous(limits = c(0,10), n.breaks = 11)
     p <- p + theme_sjplot2()
     p
     save_plot(here("Plots", "dis-dist.png"))
```

```{r Regression Analysis}
     m1 <- zeroinfl(num.dis ~ Educ + Age + Gender + 
                         Race + polviews + year |
                         Age + Age2 + Race + year, 
                         data = reg.dat
               )
     tab_model(m1, transform = NULL, file = here("Tabs", "zinf-reg.html"))
```