### 2024-01-21 analysis.qmd
updated analysis file

---
  title: "Sociological Science Categorical Tolerance Follow-up"
  author: "Omar Lizardo"
  date: "`r Sys.Date()`"
---

# Setup
```{r setup, include=FALSE}
     knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
     library(boot)
     library(conflicted)
     library(here)
     library(googlesheets4)
     library(ggplot2)
     library(pscl)
     library(sjPlot)
     library(tidyverse)
     library(webshot)
     source(here("Functions", "dat.demog.R"))
     source(here("Functions", "dat.genres.R"))
```

# Data Wrangling
```{r Getting and wrangling data}
    gs4_deauth()
    dat <- read_sheet("https://docs.google.com/spreadsheets/d/1r6IcGtlTOaAFHxJYgnuGM6IEiJhaIoDBhxEfRIhZEeI/edit?usp=sharing") 
    dat <- dat[-c(1), ] #getting rid of first row
    #working.dat <- dat.wrang(dat)
```

```{r Data Wrangling Genres}
    dat.dis <- dat.genres(dat)$dat.g %>%
          mutate(across(Classical:Metal, function(x) {if_else(x < 3, 1, 0)})) %>% 
          rowwise() %>% 
          mutate(num.dis = sum(c_across(Classical:Metal), na.rm = T)) %>% 
          ungroup() %>% 
          mutate(bin.dis = if_else(num.dis == 0, 1, 0))
    head(dat.dis)
         reg.dat <- dat.demog(dat) %>% 
          left_join(dat.dis)
```

```{r Univariate Distribution of Dislikes Plot}
     d <- round(table(dat.dis$num.dis)/nrow(dat.dis), 2)[1:11]
     plot.dat <- data.frame(cbind(count = 0:10, prop = d))
     p <- ggplot(data = plot.dat, aes(x = count, y = prop)) 
     p <- p + geom_vline(xintercept = 1, color = "red", linetype = 2, linewidth = .75)
     p <- p + geom_line(color = "blue")
     p <- p + geom_point(size = 4, color = c("red", rep("blue", 10)))
     p <- p + scale_x_continuous(limits = c(0,10), n.breaks = 11)
     p <- p + theme_sjplot2() + labs(x = "N. of Dislikes", y = "")
     p
     save_plot(here("Plots", "dis-dist.png"))
```

```{r Bivariate Plots (Race)}
      plot.dat <- reg.dat %>% 
        na.omit() %>% 
        group_by(Age) %>% 
        summarize(y = sum(bin.dis)) %>% 
        mutate(t = table(reg.dat$Age)) %>% 
        mutate(prop = round(y / t, 3))
      plot.dat$Age <- factor(plot.dat$Age, 
                             labels = c("18 - 19", "20 - 24",
                                        "25 - 29", "30 - 34", "35 - 39", "40 - 44",
                                        "45 - 49", "50 - 54", "55 - 59", "60 - 64",
                                        "65 - 69", "70+"))
      p <- ggplot(data = plot.dat, aes(y = prop, x = Age)) 
      p <- p + geom_bar(stat = "identity", fill = "blue", color = "white")
      p <- p + geom_hline(yintercept = mean(reg.dat$bin.dis, na.rm = T), color = "red", linetype = 2, linewidth = 1)
      p <- p + coord_flip()
      p <- p + theme_sjplot2() + labs(x = "", y = "")
      p
     save_plot(here("Plots", "cat-tol-age.png"))

      plot.dat <- reg.dat %>% 
        na.omit() %>% 
        group_by(Race) %>% 
        summarize(y = sum(bin.dis)) %>% 
        mutate(t = table(reg.dat$Race)) %>% 
        mutate(prop = round(y / t, 3))
      p <- ggplot(data = plot.dat, aes(y = prop, x = Race)) 
      p <- p + geom_bar(stat = "identity", fill = "blue", color = "white")
      p <- p + geom_hline(yintercept = mean(reg.dat$bin.dis, na.rm = T), color = "red", linetype = 2, linewidth = 1)
      p <- p + coord_flip()
      p <- p + theme_sjplot2() + labs(x = "", y = "")
      p
     save_plot(here("Plots", "cat-tol-race.png"))

      plot.dat <- reg.dat %>% 
        na.omit() %>% 
        group_by(Gender) %>% 
        summarize(y = sum(bin.dis)) %>% 
        mutate(t = table(reg.dat$Gender)) %>% 
        mutate(prop = round(y / t, 3))
      p <- ggplot(data = plot.dat, aes(y = prop, x = Gender)) 
      p <- p + geom_bar(stat = "identity", fill = "blue", color = "white")
      p <- p + geom_hline(yintercept = mean(reg.dat$bin.dis, na.rm = T), color = "red", linetype = 2, linewidth = 1)
      p <- p + scale_y_continuous(limits = c(0, .3))
      p <- p + coord_flip()
      p <- p + theme_sjplot2() + labs(x = "", y = "")
      p
     save_plot(here("Plots", "cat-tol-gender.png"))

      plot.dat <- reg.dat %>% 
        na.omit() %>% 
        group_by(Educ) %>% 
        summarize(y = sum(bin.dis)) %>% 
        mutate(t = table(reg.dat$Educ)) %>% 
        mutate(prop = round(y / t, 3))
      plot.dat$Educ <- factor(plot.dat$Educ, 
                             labels = c("High School", "2yr. Degree", "Bachelor's", "PhD/Prof."))
      p <- ggplot(data = plot.dat, aes(y = prop, x = Educ)) 
      p <- p + geom_bar(stat = "identity", fill = "blue", color = "white")
      p <- p + geom_hline(yintercept = mean(reg.dat$bin.dis, na.rm = T), color = "red", linetype = 2, linewidth = 1)
      p <- p + coord_flip()
      p <- p + theme_sjplot2() + labs(x = "", y = "")
      p
     save_plot(here("Plots", "cat-tol-educ.png"))

      plot.dat <- reg.dat %>% 
        na.omit() %>% 
        group_by(polviews) %>% 
        summarize(y = sum(bin.dis)) %>% 
        mutate(t = table(reg.dat$polviews)) %>% 
        mutate(prop = round(y / t, 3))
      plot.dat$polviews <- factor(plot.dat$polviews, 
                             labels = c("Extremely liberal", "Liberal",
                              "Slightly liberal", "Moderate",
                              "Slightly conservative", "Conservative"))
      p <- ggplot(data = plot.dat, aes(y = prop, x = polviews)) 
      p <- p + geom_bar(stat = "identity", fill = "blue", color = "white")
      p <- p + geom_hline(yintercept = mean(reg.dat$bin.dis, na.rm = T), color = "red", linetype = 2, linewidth = 1)
      p <- p + coord_flip()
      p <- p + theme_sjplot2() + labs(x = "", y = "")
      p
     save_plot(here("Plots", "cat-tol-pol.png"))
```

```{r Regression Analysis}
     m1 <- zeroinfl(num.dis ~ Educ + Age + Gender + 
                         Race + polviews + year |
                         Educ + Age + Age2 + Gender + Race + polviews + year, 
                         data = reg.dat
               )
     tab_model(m1, transform = NULL, file = here("Tabs", "zinf-reg.html"))
```