### 2024-01-21 analysis.qmd
Latest regression table

---
  title: "Sociological Science Categorical Tolerance Follow-up"
  author: "Omar Lizardo"
  date: "`r Sys.Date()`"
---

# Setup
```{r setup, include=FALSE}
     knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
     library(boot)
     library(conflicted)
     library(here)
     library(googlesheets4)
     library(ggplot2)
     library(pscl)
     library(sjPlot)
     library(tidyverse)
     library(webshot)
     source(here("Functions", "dat.demog.R"))
     source(here("Functions", "dat.genres.R"))
```

# Data Wrangling
```{r Getting and wrangling data}
    gs4_deauth()
    dat <- read_sheet("https://docs.google.com/spreadsheets/d/1r6IcGtlTOaAFHxJYgnuGM6IEiJhaIoDBhxEfRIhZEeI/edit?usp=sharing") 
    dat <- dat[-c(1), ] #getting rid of first row
```

```{r Data Wrangling Genres}
    dat.dis <- dat.genres(dat)$dat.g %>%
          mutate(across(Classical:Metal, function(x) {if_else(x < 3, 1, 0)})) %>% 
          rowwise() %>% 
          mutate(num.dis = sum(c_across(Classical:Metal), na.rm = T)) %>% 
          ungroup() %>% 
          mutate(bin.dis = if_else(num.dis == 0, 1, 0))
    head(dat.dis)
         reg.dat <- dat.demog(dat) %>% 
          left_join(dat.dis) %>% 
          mutate(Man = if_else(Gender == "Man", 1, 0)) %>% 
          mutate(Moderate = if_else(polviews < 3 | polviews > 5, 1, 0)) %>% 
          mutate(Black = if_else(Race == "Black", 1, 0)) %>% 
          mutate(Age_30_to_49 = if_else(Age > 3 & Age < 8, 1, 0)) %>% 
          rename(Survey_Year = year, Political_Ideology = polviews, Education = Educ) %>% 
          mutate(Survey_Year = factor(Survey_Year, labels = c("2021", "2023", "2023")))
```

```{r Univariate Distribution of Dislikes Plot}
     d <- round(table(dat.dis$num.dis)/nrow(dat.dis), 2)[1:11]
     plot.dat <- data.frame(cbind(count = 0:10, prop = d))
     p <- ggplot(data = plot.dat, aes(x = count, y = prop)) 
     p <- p + geom_vline(xintercept = 1, color = "red", linetype = 2, linewidth = .75)
     p <- p + geom_line(color = "blue")
     p <- p + geom_point(size = 4, color = c("red", rep("blue", 10)))
     p <- p + scale_x_continuous(limits = c(0,10), n.breaks = 11)
     p <- p + scale_y_continuous(limits = c(0, .25))
     p <- p + theme_sjplot2() + labs(x = "N. of Dislikes", y = "")
     p
     save_plot(here("Plots", "dis-dist.png"), 
               height = 15, width = 20)
```

```{r Univariate Distribution of Dislikes Plot (by year)}
    a <- dplyr::filter(reg.dat, year == "06 21")
    b <- dplyr::filter(reg.dat, year != "06 21")
    c <- round(table(a$num.dis)/nrow(a), 2)[1:11]
    d <- round(table(b$num.dis)/nrow(b), 2)[1:11]
    plot.dat <- data.frame(cbind(count = 0:10, y.2021 = c, y.2023 = d)) %>% 
      pivot_longer(
      cols = c(2:3),
      names_to = "Year",
      values_to = "Prop"
    ) %>% 
      mutate(Year = factor(Year, labels = c("2021", "2023"))) %>% 
      mutate(count = as.integer(count))
    head(plot.dat)

     p <- ggplot(data = plot.dat, 
                 aes(x = count, y = Prop, color = Year, group = Year)) 
     p <- p + geom_vline(xintercept = 1, color = "red", linetype = 2, linewidth = .75)
     p <- p + geom_line(linewidth = 1) + geom_point(size = 5)
     p <- p + scale_x_continuous(limits = c(0,10), n.breaks = 11)
     p <- p + scale_y_continuous(limits = c(0, .3))
     p <- p + theme_sjplot2() + labs(x = "N. of Dislikes", y = "")
     p <- p + scale_color_manual(values = c("red", "blue"))
     p <- p + theme(legend.position = "top")
     p
     save_plot(here("Plots", "dis-dist-year.png"), 
               height = 15, width = 20)
```

```{r Bar Plot Function}
    make.bar <-function(w, x, y, plotname, m){
      col_names <- names(w)
      col_name_x <- col_names[[x]]
      col_name_y <- col_names[[y]]

      p <- ggplot(data = w,  aes(x=.data[[col_name_x]], y=.data[[col_name_y]])) 
      p <- p + geom_bar(stat = "identity", fill = "blue", color = "white")
      p <- p + geom_vline(xintercept = m, color = "red", linetype = 2, linewidth = 1)
      #p <- p + coord_flip()
      p <- p + theme_sjplot2() + labs(x = "", y = "")
      p
     save_plot(here("Plots", plotname), height = 10, width = 20)
    }

```

```{r Bivariate Bar Plots (Categorical Tolerance)}
      a <- reg.dat$bin.dis
      plot.dat <- reg.dat %>% 
        na.omit() %>% 
        group_by(Age) %>% 
        summarize(y = sum(bin.dis)) %>% 
        mutate(t = table(reg.dat$Age)) %>% 
        mutate(prop = round(y / t, 3)) %>% 
        dplyr::select(c("Age", "prop"))
      plot.dat$Age <- factor(plot.dat$Age, 
                             labels = c("18 - 19", "20 - 24",
                                        "25 - 29", "30 - 34", "35 - 39", "40 - 44",
                                        "45 - 49", "50 - 54", "55 - 59", "60 - 64",
                                        "65 - 69", "70+"))
      make.bar(plot.dat, x = 2, y = 1, "cat-tol-age.png", m = mean(a, na.rm = T))

      plot.dat <- reg.dat %>% 
        na.omit() %>% 
        group_by(Race) %>% 
        summarize(y = sum(bin.dis)) %>% 
        mutate(t = table(reg.dat$Race)) %>% 
        mutate(prop = round(y / t, 3)) %>% 
        dplyr::select(c("Race", "prop"))

      make.bar(plot.dat, x = 2, y = 1, "cat-tol-race.png", m = mean(a, na.rm = T))

      plot.dat <- reg.dat %>% 
        na.omit() %>% 
        group_by(Gender) %>% 
        summarize(y = sum(bin.dis)) %>% 
        mutate(t = table(reg.dat$Gender)) %>% 
        mutate(prop = round(y / t, 3)) %>% 
        dplyr::select(c("Gender", "prop"))
      make.bar(plot.dat, x = 2, y = 1, "cat-tol-gender.png", m = mean(a, na.rm = T))

      plot.dat <- reg.dat %>% 
        na.omit() %>% 
        group_by(Educ) %>% 
        summarize(y = sum(bin.dis)) %>% 
        mutate(t = table(reg.dat$Educ)) %>% 
        mutate(prop = round(y / t, 3)) %>% 
        dplyr::select(c("Educ", "prop"))

      plot.dat$Educ <- factor(plot.dat$Educ, 
                             labels = c("High School", "2yr. Degree", "Bachelor's", "PhD/Prof."))
       make.bar(plot.dat, x = 2, y = 1, "cat-tol-educ.png", m = mean(a, na.rm = T))

      plot.dat <- reg.dat %>% 
        na.omit() %>% 
        group_by(polviews) %>% 
        summarize(y = sum(bin.dis)) %>% 
        mutate(t = table(reg.dat$polviews)) %>% 
        mutate(prop = round(y / t, 3)) %>% 
        dplyr::select(c("polviews", "prop"))

      plot.dat$polviews <- factor(plot.dat$polviews, 
                             labels = c("Extremely liberal", "Liberal",
                              "Slightly liberal", "Moderate",
                              "Slightly conservative", "Conservative"))
       make.bar(plot.dat, x = 2, y = 1, "cat-tol-pol.png", m = mean(a, na.rm = T))

```

```{r Bivariate Bar Plots (N. of Dislikes)}
      a <- reg.dat$num.dis[!reg.dat$num.dis %in% 0]
      plot.dat <- reg.dat %>% 
        na.omit() %>% 
        dplyr::filter(num.dis > 0) %>% 
        group_by(Age) %>% 
        summarize(m = mean(num.dis)) 
      plot.dat$Age <- factor(plot.dat$Age, 
                             labels = c("18 - 19", "20 - 24",
                                        "25 - 29", "30 - 34", "35 - 39", "40 - 44",
                                        "45 - 49", "50 - 54", "55 - 59", "60 - 64",
                                        "65 - 69", "70+"))
      make.bar(plot.dat, x = 2, y = 1, "num-dis-age.png", m = mean(a, na.rm = T))

      plot.dat <- reg.dat %>% 
        na.omit() %>% 
        dplyr::filter(num.dis > 0) %>% 
        group_by(Race) %>% 
        summarize(m = mean(num.dis)) 
      make.bar(plot.dat, x = 2, y = 1, "num-dis-race.png", m = mean(a, na.rm = T))

      plot.dat <- reg.dat %>% 
        na.omit() %>% 
        dplyr::filter(num.dis > 0) %>% 
        group_by(Gender) %>% 
        summarize(m = mean(num.dis)) 
      make.bar(plot.dat, x = 2, y = 1, "num-dis-gender.png", m = mean(a, na.rm = T))

      plot.dat <- reg.dat %>% 
        na.omit() %>% 
        dplyr::filter(num.dis > 0) %>% 
        group_by(Educ) %>% 
        summarize(m = mean(num.dis)) 
      plot.dat$Educ <- factor(plot.dat$Educ, 
                             labels = c("High School", "2yr. Degree", 
                                        "Bachelor's", "PhD/Prof."))
      make.bar(plot.dat, x = 2, y = 1, "num-dis-educ.png", m = mean(a, na.rm = T))

      plot.dat <- reg.dat %>% 
        na.omit() %>% 
        dplyr::filter(num.dis > 0) %>% 
        group_by(polviews) %>% 
        summarize(m = mean(num.dis)) 
      plot.dat$polviews <- factor(plot.dat$polviews, 
                             labels = c("Extremely liberal", "Liberal",
                              "Slightly liberal", "Moderate",
                              "Slightly conservative", "Conservative"))
      make.bar(plot.dat, x = 2, y = 1, "num-dis-pol.png", m = mean(a, na.rm = T))
```

```{r Regression Analysis}
     m1 <- zeroinfl(num.dis ~ 
                      Age  + Political_Ideology + Race |
                      Education + Age_30_to_49 + Man + Moderate + Black + Survey_Year,
                      data = reg.dat)
     m2 <- zeroinfl(num.dis ~ 
                      Age  + Political_Ideology + Race |
                      Education*Survey_Year + Age_30_to_49*Survey_Year + 
                      Man*Survey_Year + Moderate*Survey_Year + Black*Survey_Year,
                      data = reg.dat)
     tab_model(m1, transform = NULL, file = here("Tabs", "zinf-reg.html"), 
               auto.label = TRUE, dv.labels = "", 
               show.zeroinf = TRUE, show.aic = TRUE, show.loglik = TRUE,
               show.ci = FALSE)
     webshot(here("Tabs", "zinf-reg.html"), here("Tabs", "zinf-reg.png"))
```