### 2024-05-20 duke-presentation.qmd

---
title: "The Forward March of Categorical Tolerance in the United States?"
author: "Omar Lizardo"
institute: "*University of California, Los Angeles*"
date: "`r Sys.Date()`"
slide-number: true
show-slide-number: print
slide-level: 2
format:
  revealjs:
    incremental: false 
    theme: simple 
    smaller: true
    footer: "https://github.com/olizardo/sociological-science-categorical-tolerance-followup"
---

## Categorical Tolerance As a Social Fact

- What is **categorical tolerance**?
  - Refusal to report dislikes in survey data on cultural (usually musical) tastes
  - Anchoring at "neutral" or "mixed feelings"
- Riff on Bryson (1996) who used musical dislikes as an extension of the notion of political tolerance into the cultural realm

:::aside
Bryson, B. (1996). " Anything but heavy metal": Symbolic exclusion and musical dislikes. *American Sociological Review*, 884-899.
:::

## Categorical Tolerance As a Social Fact

- What is **categorical tolerance**?
  - Refusal to report dislikes in survey data on cultural (usually musical) tastes
  - Anchoring at "neutral" or "mixed feelings"
- Riff on Bryson (1996) who used musical dislikes as an extension of the notion of political tolerance into the cultural realm

- More prosaically:  
  - People with zero dislikes in a cultural taste survey (categorical tolerants)
  - Excess zeros in the count distribution dislikes at the population level
  - Structural zeros in a hurdle count model

:::aside
Bryson, B. (1996). " Anything but heavy metal": Symbolic exclusion and musical dislikes. *American Sociological Review*, 884-899.
:::

## Categorical Tolerance in the U.S.
![From: Lizardo, O., & Skiles, S. (2016). The end of symbolic exclusion? The rise of “categorical tolerance” in the musical tastes of Americans: 1993–2012. *Sociological Science*, 3, 85-108.](Plots/Figure1.png)

## Subsequent Work on Categorical Tolerance
::::{.columns}
::: {.column width="30%"}
![](lindbloom.png)
:::
::: {.column width="70%"}
- One comparable study by Lindbloom (2022)
- Two wave cross-section: 2007 - 2018
- Finnish Context
- 13 Musical Genres
  - Some, but not perfect overlap with Lizardo & Skiles (2016)

:::
::::

:::aside
Lindblom, T. (2022). Growing openness or creeping intolerance? Cultural taste orientations and tolerant social attitudes in Finland, 2007–2018. *Poetics*, 93, 101663.
:::

## Subsequent Work on Categorical Tolerance
::::{.columns}
::: {.column width="30%"}
![](lindbloom.png)
:::
::: {.column width="70%"}
- Findings: 
  - Categorical tolerance (CT) increases over time:
    - 14% versus 20%
  - No connection between CT and political and social attitudes
  - Only consistent predictor of CT is age:
    - Younger more likely to belong to CT group
    - Similar to Lizardo & Skiles (2016)

:::
::::

:::aside
Lindblom, T. (2022). Growing openness or creeping intolerance? Cultural taste orientations and tolerant social attitudes in Finland, 2007–2018. *Poetics*, 93, 101663.
:::

## Categorical Tolerance in the U.S. Today
:::: {.columns}
::: {.column width="50%"}
{{< include Tabs/desc-tab1.html >}}
:::
::: {.column width="50%"}
{{< include Tabs/desc-tab2.html >}}
:::
::::

## Categorical Tolerance in the U.S. Today
:::: {.columns}
::: {.column width="45%"}
- Unweighted distribution over-estimates categorical tolerance effect
- Weighted distribution similar to that observed in 2012 data
  - 2012 SSI Data: 
    - 16% Categorical Tolerants
  - 2021/2023 Prolific Data: 
    - 20% Categorical Tolerants (Unweighted)
    - 14% Categorical Tolerants (Weighted)
:::
::: {.column width="55%"}
![Data from 2021/2023 (Pooled)](Plots/uni-dist-by-weighted.png)
:::
::::

## Differences in Two Data Collection Waves
:::: {.columns}
::: {.column width="45%"}
- Most of the categorical tolerance over-estimation comes from smaller data collection wave in 2021
  - Middle of Covid Pandemic
- Categorical Tolerance 2021:
  - Unweighted: 27%
  - Weighted: 9%
- Categorical Tolerance 2023:
  - Unweighted: 16%
  - Weighted: 17% 
:::
::: {.column width="55%"}
![Data from 2021/2023](Plots/uni-dist-by-weighted-by-year.png)
:::
::::

## Socio-Demographic Correlates 
### Categorical Tolerance and Age
:::: {.columns}
::: {.column width="65%"}
![](Plots/cat-tol-age.png)
:::
::: {.column width="35%"}
- Young people over-represented among categorical tolerants
- Old people under-represented
- Consistent with Lizardo & Skiles (2016)
:::
::::

## Socio-Demographic Correlates 
### Categorical Tolerance and Race
:::: {.columns}
::: {.column width="65%"}
![](Plots/cat-tol-race.png)
:::
::: {.column width="35%"}
- White people under-represented among categorical tolerants
  - 58% versus 40%

- Black people over-represented 
    - 13% versus 23% 
- People identifying as "multiracial" also over-represented
  - Consistent with Lizardo & Skiles (2016)
:::
::::

## Socio-Demographic Correlates 
### Categorical Tolerance and Gender
:::: {.columns}
::: {.column width="65%"}
![](Plots/cat-tol-gender.png)
:::
::: {.column width="35%"}
- Strong gender effect 
- Women under-represented among categorical tolerants
- Men over-represented
:::
::::

## Socio-Demographic Correlates 
### Categorical Tolerance and Education
:::: {.columns}
::: {.column width="65%"}
![](Plots/cat-tol-educ.png)
:::
::: {.column width="35%"}
- Not much of an education effect on categorical tolerance
  - Consistent with Lizardo & Skiles (2016)

:::
::::

## Socio-Demographic Correlates 
### Categorical Tolerance and Political Ideology
:::: {.columns}
::: {.column width="70%"}
![](Plots/cat-tol-pol.png)
:::
::: {.column width="30%"}
- Not much of a Political Ideology effect
  - Not considered by Lizardo & Skiles (2016)
  - SSI 2012 Data lacked information on Political Ideology
:::
::::

## Socio-Demographic Correlates 
### Graded Tolerance and Age
:::: {.columns}
::: {.column width="65%"}
![](Plots/num-dis-age.png)
:::

::: {.column width="35%"}
- Dislikes increase with age
- Not much difference between weighted and unweighted data
  - Especially at the old end
  - No Surprise there!
:::
::::

## Socio-Demographic Correlates

### Graded Tolerance and Education
:::: {.columns}
::: {.column width="65%"}
![](Plots/num-dis-educ.png)
:::

::: {.column width="35%"}
- Curvilinear pattern (weighted data)
- Symbolic Exclusion higher at low and high end of the status ladder
- Middle status tolerance effect
- Not particularly strong
:::
::::

## Socio-Demographic Correlates 
### Graded Tolerance and Political Ideology
:::: {.columns}
::: {.column width="65%"}
![](Plots/num-dis-pol.png)
:::

::: {.column width="35%"}
- Strong ideology effect on graded tolerance
  - In contrast to categorical tolerance
- Expected direction:
  - Liberals more tolerant
  - Conservatives more likely to engage in symbolic exclusion
  - Pattern more salient in weighted data
:::
::::

## Socio-Demographic Correlates 
### Graded Tolerance and Race
:::: {.columns}
::: {.column width="65%"}
![](Plots/num-dis-race.png)
:::

::: {.column width="35%"}
- Race effect on graded exclusion
- White/non-White Contrast
- White people more likely to engage in symbolic exclusion
:::
::::

## Socio-Demographic Correlates 
### Graded Tolerance and Gender
:::: {.columns}
::: {.column width="65%"}
![](Plots/num-dis-gender.png)
:::

::: {.column width="35%"}
- Not much of a gender effect on graded exclusion
- Non/binary v. Everyone Else contrast in weighted data
:::
::::

## Zero-Inflated Poisson Regression Results 
:::: {.columns}
:::{.column width="40%"}
{{< include Tabs/zinf-reg.html >}}
:::
:::{.column width="60%"}
- Categorical Tolerance (Zero Inflation)
  - Black and Multiracial Positive Effect
    - Hispanic Not Statistically Signficant
  - Generational Effect
     - Old versus Young
  - Gender Effect
     - Women Less Likely to be Categorically Tolerant
  - Period Effect
     - Covid Depressed Categorical Tolerance 
     - Effect Evident Once Sample is Re-weighted

:::
::::

## Zero-Inflated Poisson Regression Results 
:::: {.columns}
:::{.column width="40%"}
{{< include Tabs/zinf-reg.html >}}
:::
:::{.column width="60%"}
- Graded Exclusion (Count Model)
  - Age and Political Ideology Effects in Expected Direction
    - Older, more conservative more exclusion
  - Education Effect (Weighted Data)
    - More Symbolic Exclusion at the Extreme Low and High Ends
  - Race Effect in Expected Direction
    - White people more likely to dislike

:::
::::

## Conclusions
- Categorical Tolerance continues to be a social fact!
  - Highlights value of replication of previous work using newer/different data
- Similarity in results with previous work in the U.S. re-assuring
  - CTs are non-white and young today as they were in 2012
- Similarity with results from a different national contemporary national case suggest generic not case-specific mechanisms
  - Especially with regard to the cohort/age effect
  - Race effect could be specific to the U.S. Context
- Need work using other methodological strategies (interview, experimental, computational) to crack the puzzle

## ZIP Regression Appendix

{{< include Tabs/zinf-reg-appendix.html >}}