### 2025-06-07 analysis-old.qmd
buncha stuff

---
  title: "Sociological Science Categorical Tolerance Follow-up"
  author: "Omar Lizardo"
  date: "`r Sys.Date()`"
---

# Setup
```{r setup, include=FALSE}
     knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
     library(anesrake)
     library(boot)
     library(conflicted)
     library(here)
     library(googlesheets4)
     library(ggplot2)
     library(kableExtra)
     library(pscl)
     library(questionr)
     library(sjPlot)
     library(tidyverse)
     library(webshot)
     source(here("Functions", "dat.demog.R"))
     source(here("Functions", "dat.genres.R"))
     conflicted::conflicts_prefer(questionr::wtd.table)
     conflicted::conflicts_prefer(dplyr::summarize)
```

# Wrangling
```{r Getting and wrangling data}
    gs4_deauth()
    dat <- read_sheet("https://docs.google.com/spreadsheets/d/1r6IcGtlTOaAFHxJYgnuGM6IEiJhaIoDBhxEfRIhZEeI/edit?usp=sharing") 
    dat <- dat[-c(1), ] #getting rid of first row
```

```{r Data Wrangling}
    dat.dis <- dat.genres(dat)$dat.g %>%
          mutate(across(Classical:Metal, function(x) {if_else(x < 3, 1, 0)})) %>% 
          rowwise() %>% 
          mutate(num.dis = sum(c_across(Classical:Metal), na.rm = T)) %>% 
          ungroup() %>% 
          mutate(bin.dis = if_else(num.dis == 0, 1, 0))
    reg.dat <- dat.demog(dat) %>% 
          left_join(dat.dis) %>% 
          rename(Survey_Year = year, Political_Ideology = polviews, Education = Educ) %>%
          mutate(Black = if_else(Race == "Black", 1, 0)) %>% 
          mutate(Hisp = if_else(Race == "Hisp.", 1, 0)) %>% 
          mutate(Asian = if_else(Race == "Asian", 1, 0)) %>% 
          mutate(White = if_else(Race == "White", 1, 0)) %>%
          mutate(Mult = if_else(Race == "Mult.", 1, 0)) %>% 
          mutate(Native = if_else(Race == "Native/Other", 1, 0)) %>% 
          mutate(Woman = if_else(Gender == "Woman", 1, 0)) %>% 
          mutate(Age.n = as.numeric(Age)) %>% 
          mutate(Young = if_else(Age.n > 1 & Age.n < 5, 1, 0)) %>% 
          mutate(Old = if_else(Age.n > 9, 1, 0)) %>% 
          mutate(Educ.n = as.numeric(Education)) %>% 
          mutate(Low.Educ = if_else(Educ.n == 1, 1, 0)) %>% 
          mutate(Mid.Educ = if_else(Educ.n == 2 | Educ.n == 3, 1, 0)) %>% 
          mutate(Hi.Educ = if_else(Educ.n == 4, 1, 0)) %>% 
          mutate(polviews.n = as.numeric(Political_Ideology)) %>% 
          mutate(Moderate = if_else(polviews.n == 3 | polviews.n == 4, 1, 0)) %>% 
          mutate(Survey_Year = factor(Survey_Year, labels = c("2021", "2023", "2023"))) %>% 
          mutate(Age = factor(Age, 
                             labels = c("18 - 19", "20 - 24",
                                        "25 - 29", "30 - 34", "35 - 39", "40 - 44",
                                        "45 - 49", "50 - 54", "55 - 59", "60 - 64",
                                        "65 - 69", "70+"))) %>% 
         mutate(Education = factor(Education, labels = c("High School", "Associate", "Bachelor", "Grad/Prof."))) %>% 
         mutate(Political_Ideology = factor(Political_Ideology, 
                             labels = c("Extremely liberal", "Liberal",
                              "Slightly liberal", "Moderate",
                              "Slightly conservative", "Conservative")))
```

# Weighting
```{r Target Proportions}
    Race <- c(0.589, 0.136, 0.063, 0.191, 0.03, 0.016)
    names(Race) <- c("White", "Black", "Asian", "Hisp.", "Mult.", "Native/Other")
    Age <- c(0.034, 0.084, 0.089, 0.088, 0.086, 0.080, 0.081, 0.082, 0.086, 0.082, 0.069, 0.138)
    names(Age) <- c("18 - 19", "20 - 24", "25 - 29", "30 - 34", "35 - 39", "40 - 44", "45 - 49", "50 - 54", "55 - 59", "60 - 64", "65 - 69", "70+")
    Education <- c(0.58, 0.09, 0.21, 0.13)
    names(Education) <- c("High School", "Associate", "Bachelor", "Grad/Prof.")
    Gender <- c(.495, .495, .01)
    names(Gender) <- levels(reg.dat$Gender)
    target <- list(Race = Race, Education = Education, Age = Age, Gender = Gender)
```

```{r Creating weights via raking}
    rake.dat <- as.data.frame(reg.dat)
    raking <- anesrake::anesrake(target, rake.dat, rake.dat$id, 
                                 maxit = 5000, pctlim = 0.025,
                                 choosemethod = "average")
    reg.dat$weight <- raking$weightvec
    reg.dat.21 <- dplyr::filter(reg.dat, Survey_Year == "2021")
    reg.dat.23 <- dplyr::filter(reg.dat, Survey_Year == "2023")
```

# Descriptive Tables
```{r Descriptive sociodemographics table}
     prop <- function(x, labs) {
          f <- table(as.numeric(x))
          p <- round(f/sum(f), 3) * 100
          w.f <- wtd.table(as.numeric(x), weights = reg.dat$weight)
          w.p <- round(w.f/sum(w.f), 3) * 100
          rownames(f) <- labs
          z <- as.matrix(cbind(f, p, w.p)) 
          colnames(z) <- c("N. of Cases", "Per.", "Weighted Per.")
     return(z)
     }

     gender.tab <- prop(reg.dat$Gender, 
                        c("Man", "Woman", "Other/Non-binary"))
     educ.tab <- prop(reg.dat$Education, 
                      c("High School", "Associate", "Bachelors", "PhD/Prof."))
     age.tab <- prop(reg.dat$Age, c("18 - 19", "20 - 24", "25 -29", 
                                         "30 - 34", "35 - 39", "40 - 44", 
                                         "45 - 49", "50 - 54", "55 - 59", 
                                         "60 - 64", "65 - 69", "70+"))
     pol.tab <- prop(reg.dat$Political_Ideology,
                     rownames(table(reg.dat$Political_Ideology)))
     race.tab <- prop(reg.dat$Race, rownames(table(reg.dat$Race)))
     year.tab <- prop(reg.dat$Survey_Year, c("2021", "2023"))
     tab1 <- rbind(educ.tab, race.tab, gender.tab) %>% 
          kbl(caption = "", format = "html", align = "c") %>% 
          kable_classic_2(full_width = F, 
                          html_font = "Garamond", font_size = 20) %>% 
          pack_rows("Education", 1, 4) %>% 
          pack_rows("Ethnic/Racial I.D.", 5, 10) %>% 
          pack_rows("Gender I.D.", 11, 13) %>% 
          save_kable(here("Tabs", "desc-tab1.html"))

     tab2 <- rbind(age.tab, year.tab) %>% 
          kbl(caption = "", format = "html", align = "c") %>% 
          kable_classic_2(full_width = F, 
                          html_font = "Garamond", font_size = 20) %>% 
          pack_rows("Age Groups", 1, 12) %>% 
          pack_rows("Survey Year", 13, 14) %>%   
          save_kable(here("Tabs", "desc-tab2.html"))
```

```{r}
  dat.g <- reg.dat[, c("id", "Classical", "Country", "Disco",
                            "Jazz", "Musicals", "Pop", "R_and_B",
                            "Rap", "Reggae", "Rock", "EDM", "Metal")] %>% 
    pivot_longer(
      cols = c(2:13),
      names_to = "genre",
      values_to = "Dislike"
    ) 
  t <- round(table(dat.g$genre, dat.g$Dislike)/nrow(reg.dat), 2) * 100
  t.w <- left_join(dat.g, reg.dat[, c("id", "weight")])
  t.w <- round(wtd.table(t.w$genre, t.w$Dislike, weights = t.w$weight)/nrow(reg.dat), 2) * 100
  t <- data.frame(d = t[, 2], wd = t.w[, 2]) %>% 
    kbl(caption = "", format = "html", align = "c", 
        col.names = c("Unweighted", "Weighted")) %>% 
          kable_classic_2(full_width = F, 
          html_font = "Garamond", font_size = 20) %>% 
          pack_rows("% Disliked", 1, 12) %>% 
          save_kable(here("Tabs", "desc-tab3.html"))

```

# Plot Functions
```{r Plot Data Wrangling}
      make.plot.dat <- function(x) {
        l <- length(table(x))
        p.u <- wtd.table(x, weights = reg.dat$weight)/sum(wtd.table(x, weights = reg.dat$weight))
        p.w <- wtd.table(x, reg.dat$bin.dis, weights = reg.dat$weight)
        p.w <- p.w[, 2]/sum(p.w[, 2])
        plot.dat <- data.frame(cbind(p.u = round(p.u, 3), 
                                     p.w = round(p.w, 3), 
                                     x.var = names(p.u))) %>% 
        pivot_longer(
        cols = c(1:2),
        names_to = "Weighted",
        values_to = "Prop"
        ) %>% 
        mutate(Weighted = factor(Weighted, labels = c("Sample Prop. (Weighted)", "Obs. Cat. Tol. (Weighted)"))) %>% 
        mutate(Prop = as.numeric(Prop)) 
      }

    make.plot.dat.n <- function(x) {
        w.dat <- reg.dat %>% 
          dplyr::filter(num.dis > 0) %>% 
          group_by(.data[[x]]) %>% 
          summarize(w.m = weighted.mean(num.dis, weight)) 
        plot.dat <- reg.dat %>% 
          dplyr::filter(num.dis > 0) %>% 
          group_by(.data[[x]]) %>% 
          summarize(u.m = mean(num.dis)) %>% 
          left_join(w.dat) %>% 
          pivot_longer(
          cols = c(2:3),
          names_to = "Weighted",
          values_to = "n.dis"
          ) %>% 
          mutate(Weighted = factor(Weighted, labels = c("Unweighted", "Weighted"))) %>% 
          rename(x.var = 1) %>% 
          na.omit()
          }
```

```{r Plot Making}
    bivar.line.plot <- function(x, alpha = .2) {
      dat <- make.plot.dat(x)
      l <- length(unique(dat$x.var))
      o <- levels(x)
      p <- ggplot(data = dat, aes(x = factor(x.var, levels = o), y = Prop, 
                                       color = Weighted, group = Weighted))
      p <- p + geom_line(alpha = c(rep(alpha, l), rep(1, l))) 
      p <- p + geom_point(size = 3, alpha = rep(c(alpha, 1), l))
      p <- p + theme_minimal() + labs(x = "", y = "")
      p <- p + theme(legend.position = "bottom", 
                     legend.title = element_blank(),
                     axis.line.x = element_blank(), 
                     axis.line.y = element_blank(),
                     legend.text = element_text(size = 10),
                     axis.text.y = element_text(size = 12),
                     axis.text.x = element_text(size = 12))
      p <- p + scale_color_manual(values = c("blue", "blue"))
      p <- p + guides(color = guide_legend (override.aes = list(alpha = c(alpha, 1))))
      p
    }

    bivar.line.plot.d <- function(x, alpha = .2) {
      dat <- make.plot.dat.n(x)
      l <- length(unique(dat$x.var))
      p <- ggplot(data = dat, aes(x = x.var, y = n.dis, 
                                       color = Weighted, group = Weighted))
      p <- p + geom_line(alpha = c(rep(alpha, l), rep(1, l))) 
      p <- p + geom_point(size = 3, alpha = rep(c(alpha, 1), l))
      p <- p + theme_minimal() + labs(x = "", y = "Mean N. of Dislikes (Excluding Categorical Tolerants)")
      p <- p + theme(legend.position = "bottom", 
                     legend.title = element_blank(),
                     axis.line.x = element_blank(), 
                     axis.line.y = element_blank(),
                     axis.text.x = element_text(size = 12))
      p <- p + scale_color_manual(values = c("blue", "blue"))
      p <- p + guides(color = guide_legend (override.aes = list(alpha = c(alpha, 1))))
      p <- p + ylim(2.5, 4.75)
      p
    }

    bivar.bar.plot <- function(x, alpha = .1) {
      dat <- make.plot.dat(x)
      l <- length(unique(dat$x.var))
      p <- ggplot(data = dat, aes(x = x.var, y = Prop, 
                                       fill = Weighted, 
                                       group = Weighted))
      p <- p + geom_col(position = "dodge", alpha = rep(c(alpha, 1), l), color = "white")
      p <- p + theme_minimal() + labs(x = "", y = "")
      p <- p + theme(legend.position = "bottom", 
                     legend.title = element_blank(),
                     legend.text = element_text(size = 12),
                     axis.text.y = element_text(size = 12),
                     axis.text.x = element_text(size = 12))
      p <- p + scale_fill_manual(values = c("blue", "blue")) + coord_flip()
      p <- p + guides(fill = guide_legend (override.aes = list(alpha = c(alpha, 1))))
    p
    }

    bivar.bar.plot.d <- function(x, alpha = .1) {
      dat <- make.plot.dat.n(x)
      l <- length(unique(dat$x.var))
      p <- ggplot(data = dat, aes(x = x.var, y = n.dis, 
                                       fill = Weighted, 
                                       group = Weighted))
      p <- p + geom_col(position = "dodge", alpha = rep(c(alpha, 1), l), color = "white")
      p <- p + theme_minimal() + labs(x = "", y = "Mean N. of Dislikes (Excluding Categorical Tolerants)")
      p <- p + theme(legend.position = "bottom", 
                     legend.title = element_blank(),
                     legend.text = element_text(size = 12),
                     axis.text.y = element_text(size = 12),
                     axis.text.x = element_text(size = 12))
      p <- p + scale_fill_manual(values = c("blue", "blue")) + coord_flip()
      p <- p + guides(fill = guide_legend (override.aes = list(alpha = c(alpha, 1))))
      p
    }
```

# Plots
## Univariate
### Univariate Data Wrangling
```{r Univariate Distribution of Dislikes Data}
     d.u <- round(table(reg.dat$num.dis)/nrow(reg.dat), 2)[1:11]
     d.u.21 <- round(table(reg.dat.21$num.dis)/nrow(reg.dat.21), 2)[1:11]
     d.u.23 <- round(table(reg.dat.23$num.dis)/nrow(reg.dat.23), 2)[1:11]
     d.w <- round(wtd.table(reg.dat$num.dis, weights = reg.dat$weight)/nrow(reg.dat), 2)[1:11]
     d.w.21 <- round(wtd.table(reg.dat.21$num.dis, weights = reg.dat.21$weight)/nrow(reg.dat.21), 2)[1:11]
     d.w.23 <- round(wtd.table(reg.dat.23$num.dis, weights = reg.dat.23$weight)/nrow(reg.dat.23), 2)[1:11]
    uni.plot.dat <- data.frame(cbind(count = 0:10, prop.u = d.u, prop.w = d.w)) %>% 
      pivot_longer(
        cols = c(2:3),
        names_to = "Weighted",
        values_to = "Prop"
        ) %>% 
      mutate(Weighted = factor(Weighted, labels = c("Unweighted", "Weighted"))) %>% 
      mutate(count = as.integer(count))

    uni.plot.dat.21 <- data.frame(cbind(count = 0:10, prop.u = d.u.21, prop.w = d.w.21)) %>% 
      pivot_longer(
        cols = c(2:3),
        names_to = "Weighted",
        values_to = "Prop"
        ) %>% 
      mutate(Weighted = factor(Weighted, labels = c("Unweighted", "Weighted"))) %>% 
      mutate(count = as.integer(count)) %>% 
      mutate(year = "2021")

    uni.plot.dat.23 <- data.frame(cbind(count = 0:10, prop.u = d.u.23, prop.w = d.w.23)) %>% 
      pivot_longer(
        cols = c(2:3),
        names_to = "Weighted",
        values_to = "Prop"
        ) %>% 
      mutate(Weighted = factor(Weighted, labels = c("Unweighted", "Weighted"))) %>% 
      mutate(count = as.integer(count)) %>% 
      mutate(year = "2023")
  uni.plot.dat.year <- rbind(uni.plot.dat.21, uni.plot.dat.23)
```

### Pooled data plot
```{r Univariate Distribution of Dislikes Plot (by weighted)}
     p <- ggplot(data = uni.plot.dat, aes(x = count, y = Prop, 
                                          color = Weighted, group = Weighted)) 
     p <- p + geom_vline(xintercept = 1, color = "red", 
                         linetype = 2, linewidth = .5)
     p <- p + geom_line(alpha = c(rep(.25, 11), rep(1, 11))) 
     p <- p + geom_point(size = 4, alpha = rep(c(.25, 1), 11))
     p <- p + scale_x_continuous(limits = c(0,10), n.breaks = 11)
     p <- p + theme_minimal() + labs(x = "N. of Dislikes", y = "")
     p <- p + theme(legend.text = element_text(size = 12),
                    axis.text.y = element_text(size = 12),
                    axis.text.x = element_text(size = 12),
                    axis.title.x = element_text(size = 12))
     p <- p + scale_color_manual(values = c("blue", "blue"))
     p <- p + guides(color = guide_legend (override.aes = list(alpha = c(.2, 1))))
     p <- p + theme(legend.position = "top", legend.title = element_blank())
     p
     save_plot(here("Plots", "uni-dist-by-weighted.png"), 
               height = 15, width = 15)
```

### Separate survey wave plot
```{r Univariate Distribution of Dislikes Plot (unweighted by survey year)}
     p <- ggplot(data = uni.plot.dat.year, aes(x = count, y = Prop, 
                                          color = Weighted, group = Weighted)) 
     p <- p + geom_vline(xintercept = 1, color = "red", 
                         linetype = 2, linewidth = .5)
     p <- p + geom_line(alpha = c(rep(.25, 22), rep(1, 22))) 
     p <- p + geom_point(size = 3, alpha = rep(c(.25, 1), 22))
     p <- p + scale_x_continuous(limits = c(0,10), n.breaks = 11)
     p <- p + theme_minimal() + labs(x = "N. of Dislikes", y = "")
     p <- p + scale_color_manual(values = c("blue", "blue"))
     p <- p + guides(color = guide_legend (override.aes = list(alpha = c(.2, 1))))
     p <- p + theme(legend.position = "bottom", 
                    legend.title = element_blank(),
                    strip.text = element_text(size = 14),
                    legend.text = element_text(size = 12),
                    axis.text.y = element_text(size = 12),
                    axis.text.x = element_text(size = 12),
                    axis.title.x = element_text(size = 12))
     p <- p + scale_color_manual(values = c("blue", "blue"))
     p <- p + facet_wrap(~ year, nrow = 1) + ylim(0, .3)
     p
     save_plot(here("Plots", "uni-dist-by-weighted-by-year.png"), 
               height = 18, width = 18)
```
## Bivariate (Categorical Tolerance)
```{r Bivariate Plots (Categorical Tolerance)}
    bivar.line.plot(reg.dat$Age)
    save_plot(here("Plots", "cat-tol-age.png"), height = 15, width = 20)
    bivar.line.plot(reg.dat$Education)
    save_plot(here("Plots", "cat-tol-educ.png"), height = 15, width = 20)
    bivar.line.plot(reg.dat$Political_Ideology)
    save_plot(here("Plots", "cat-tol-pol.png"), height = 16, width = 22)
    bivar.bar.plot(reg.dat$Race)
    save_plot(here("Plots", "cat-tol-race.png"), height = 15, width = 20)
    bivar.bar.plot(reg.dat$Gender)
    save_plot(here("Plots", "cat-tol-gender.png"), height = 15, width = 20)
```

## Bivariate (Graded Tolerance)
```{r Bivariate Bar Plots (N. of Dislikes)}
    bivar.line.plot.d("Age")
    save_plot(here("Plots", "num-dis-age.png"), height = 17, width = 20)
    bivar.line.plot.d("Education")
    save_plot(here("Plots", "num-dis-educ.png"), height = 12, width = 16)
    bivar.line.plot.d("Political_Ideology")
    save_plot(here("Plots", "num-dis-pol.png"), height = 16, width = 22)
    bivar.bar.plot.d("Race")
    save_plot(here("Plots", "num-dis-race.png"), height = 15, width = 20)
    bivar.bar.plot.d("Gender")
    save_plot(here("Plots", "num-dis-gender.png"), height = 15, width = 20)
```

# Regression
```{r Regression Analysis}
     m1 <- zeroinfl(
       num.dis ~  Age.n + Hi.Educ + Low.Educ + polviews.n + White | 
                  Young + Old + Woman + Black + Hisp + Mult + Survey_Year, 
       data = reg.dat
       )
     m2 <- zeroinfl(
       num.dis ~  Age.n + Hi.Educ + Low.Educ + polviews.n + White | 
                  Age.n + Young + Old + Woman + Black + Hisp + Mult + Survey_Year, 
       data = reg.dat, weights = weight
       )
     m3 <- zeroinfl(
       num.dis ~  Age.n + Hi.Educ + Low.Educ + polviews.n + White | 
                  Age.n + Young + Old + Woman + Black + Hisp + Mult,       
       data = reg.dat.21, weights = weight
       )
     m4 <- zeroinfl(
       num.dis ~  Age.n + Hi.Educ + Low.Educ + polviews.n + White | 
                  Age.n + Young + Old + Woman + Black + Hisp + Mult,       
       data = reg.dat.23, weights = weight
       )
     m5 <- zeroinfl(
       num.dis ~  Age.n + Hi.Educ + Low.Educ + polviews.n + White | 
                  Age.n + Young + Old + Woman + Black + Hisp + Mult,       
       data = reg.dat.21
       )
     m6 <- zeroinfl(
       num.dis ~  Age.n + Hi.Educ + Low.Educ + polviews.n + White | 
                  Age.n + Young + Old + Woman + Black + Hisp + Mult,       
       data = reg.dat.23
       )
     tab_model(m1, m2, transform = NULL, 
               file = here("Tabs", "zinf-reg.html"), 
               auto.label = TRUE, dv.labels = c("Unweighted", "Weighted"), 
               CSS = list(css.table = 'font-size: 12px;',
                          css.td = 'border-bottom: 0px;'), 
               show.zeroinf = TRUE, show.ci = FALSE)
     tab_model(m5, m3, m6, m4, transform = NULL, 
               file = here("Tabs", "zinf-reg-appendix.html"), 
               auto.label = TRUE, dv.labels = 
                 c("Unweighted (2021)", "Weighted (2021)", 
                   "Unweighted (2023)", "Weighted (2023)"), 
               CSS = list(css.table = 'font-size: 13px;',
                          css.td = 'border-bottom: 0px;'),               
               show.zeroinf = TRUE, show.ci = FALSE)
```