### 2025-11-15 analysis-comp.qmd
data rep

---
  title: "Sociological Science Categorical Tolerance Follow-up (Appendix)"
  author: "Omar Lizardo"
  date: "`r Sys.Date()`"
---

```{r setup, include=FALSE}
     knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
     library(conflicted)
     library(here)
     library(dplyr)
     library(tidyr)
     library(webshot)
     library(qualtRics)
     library(kableExtra)
     library(ggplot2)
     library(ggeffects)
     library(sjPlot)
     library(pscl)
     library(haven)
     conflicts_prefer(dplyr::filter)
     conflicts_prefer(dplyr::select)
     conflicts_prefer(dplyr::summarize)
     source(here("Functions", "dat.genres.R"))
```

<!-- ```{r Getting Data (Lucid and Prolific)}
    dat.25 <- fetch_survey(surveyID = "SV_80RzX3ibOsBq3Lo") |> 
        slice(-1) |> #getting rid of first row 
        filter(Q_RecaptchaScore  > 0.50) |>  #removing suspected bots
        filter(DistributionChannel != "preview") 
    dat.23 <- fetch_survey(surveyID = "SV_dd88aecMxMf0lmK") |> 
        slice(-1) |> #getting rid of first row 
        filter(Q_RecaptchaScore  > 0.50) |> #removing suspected bots
        filter(DistributionChannel != "preview")
```

```{r Wrangling Data (Lucid and Prolific)}
     dat.dis <- rbind(dat.genres(dat.25)$dat.g, dat.genres(dat.23)$dat.g) |> 
        select(-Disco) |> 
        mutate(across(Classical:Metal, function(x) {if_else(x < 3, 1, 0)})) |>
        rowwise() |>
        mutate(num.dis = sum(c_across(Classical:Metal), na.rm = TRUE)) |> 
        ungroup() 
    dat.dis.25 <- dat.genres(dat.25)$dat.g |> 
        select(-Disco) |> 
        mutate(across(Classical:Metal, function(x) {if_else(x < 3, 1, 0)})) |>
        rowwise() |>
        mutate(num.dis = sum(c_across(Classical:Metal), na.rm = TRUE)) |> 
        ungroup() 
    save(dat.dis, dat.dis.25, file = here("cint-prolific-data2.rda"))
``` -->

```{r Getting Data (SSI)}
    dat.dis.12 <- read_dta(here("SSI2012.dta")) |> 
        select("id", ends_with(c("taste")), -starts_with("none")) |> 
        select(c(1:21)) |> 
        mutate(across(classicaltaste:hvymtltaste, 
                ~case_when(. == 1 ~ 0, . == 2 ~ 1, . == 3 ~ 0, . == 4 ~ NA))
                ) |> 
        rename(
                Classical = classicaltaste, 
                Country = countrytaste,
                Jazz = jazztaste,
                Musicals = bwaysttaste,
                Pop = toppoptaste,
                R_and_B = blurbtaste,
                Rap = raphiphoptaste,
                Reggae = reggaetaste,
                Rock = croldtaste,
                Rock2 = controcktaste,
                Rock3 = indalttaste,
                EDM = danclubtaste,
                Metal = hvymtltaste
                ) |> 
        mutate(Rock = if_else(Rock == 1 | Rock2 == 1 | Rock3 == 1, 1, 0)) |> 
        select(-c(Rock2, Rock3), -ends_with("taste")) |> 
        rowwise() |>
        mutate(num.dis = sum(c_across(Classical:Metal))) |> 
        ungroup() |> 
        na.omit() 
```

```{r Univariate Distribution Data}
    load(here("cint-prolific-data2.rda"))
    dat.count.12 <- dat.dis.12 |> 
        group_by(num.dis) |>
        summarize(n = n()) |>
        mutate(Prop = n/sum(n)) |>
        rename(Count = num.dis) |>
        mutate(Count = as.factor(Count)) |>
        select(Count, Prop) |>
        data.frame() |> 
        mutate(sample = "2012")
        dat.count.12
    dat.count.25 <- dat.dis.25 |> 
        group_by(num.dis) |>
        summarize(n = n()) |>
        mutate(Prop = n/sum(n)) |>
        rename(Count = num.dis) |>
        mutate(Count = as.factor(Count)) |>
        select(Count, Prop) |>
        data.frame() |> 
        mutate(sample = "2025")
        dat.count.25
    dat.count <- dat.dis |> 
        group_by(num.dis) |>
        summarize(n = n()) |>
        mutate(Prop = n/sum(n)) |>
        rename(Count = num.dis) |>
        mutate(Count = as.factor(Count)) |>
        select(Count, Prop) |>
        data.frame() |> 
        mutate(sample = "2023/25")
        dat.count
```

```{r Testing for Equality of Distribution Across Samples}
    ks.test(dat.dis.12$num.dis, dat.dis$num.dis)
```

```{r Comparing Univariate Distribution Plot}
    dat.uni <- rbind(dat.count.12, dat.count.25)
    p <- ggplot(data = dat.uni, aes(Count, y = Prop, group = sample, color = sample))
    p <- p + geom_vline(xintercept = 2, color = "red", linetype = 2, linewidth = 1)
    p <- p + geom_line(linewidth = 1) 
    p <- p + geom_point(size = 6)
    p <- p + theme_minimal() 
    p <- p + labs(x = "Number of Dislikes", y = "")
    p <- p + theme(axis.text = element_text(size = 18),
                    axis.text.x = element_text(hjust = 0.6),
                    axis.title = element_text(size = 18),
                    legend.title = element_blank(),
                    legend.position = "top",
                    legend.text = element_text(size = 18)
                    )
    p <- p + ylim(0, 0.25)
    p <- p + scale_color_manual(values = c("lightblue", "blue"))
    p
    save_plot(here("Plots", "uni-dist-cat-tol-comp.png"), height = 15, width = 20)
```

```{r Zero Inflated Regression}
    dat.dis.12$sample = "2012"
    dat.dis.25$sample = "2025"
    dat.dis$sample = "2021/23/25"
    pooled.dat <- rbind(dat.dis.25, dat.dis.12)
    pooled.dat$sample <- factor(pooled.dat$sample)
    m1 <- zeroinfl(num.dis ~ sample | sample, data = pooled.dat)
    summary(m1)
    tab_model(m1, transform = NULL, digits = 3, prefix.labels = "none", show.stat = TRUE,
            file = here("Tabs", "zinf-reg-comp.html"), 
            auto.label = TRUE, dv.labels = c("Model 1"), 
            CSS = list(css.table = 'font-size: 12px;', css.td = 'border-bottom: 0px;'), 
            show.zeroinf = TRUE, show.ci = FALSE,
            pred.labels = c("Intercept", "Sample (2025)")
            )
    webshot(here("Tabs", "zinf-reg-comp.html"), here("Tabs", "zinf-reg-comp.png"), zoom = 4)
```

```{r Predicted Probabilities}
    pv <- predict_response(m1, "sample", type = "zi_prob")
    p <- plot(pv, ci_style = "errorbar", dot_size = 6)
    p <- p + theme_minimal()
    p <- p + theme(axis.text.y = element_text(size = 18),
                    axis.text.x = element_text(size = 24),
                    axis.title = element_blank(),
                    plot.title =  element_text(size = 24)
                    )
    p <- p + ggtitle("Predicted Zero-Inflation Probabilities") 
    p
    save_plot(here("Plots", "zip-pred-comp.png"), height = 20, width = 20)
```